---
title: "Human ATRT RNAseq data"
author: "Mamy Andrianteranagna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 9
    keep_md: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
    number_sections: true
    code_folding: hide
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, include=TRUE, cache=FALSE, error=TRUE)

```

```{r paths}

thisProject                  <- "cellOrigin_ATRT/"

# Paths
pathToRhabdoid_U830_projects <- "/bioinfo/users/mandrian/rhabdoid_U830_projects/"
pathToProgrammes             <- paste0(pathToRhabdoid_U830_projects,"programmes/")
pathToANALYSIS_PROJECTS      <- paste0(pathToRhabdoid_U830_projects,"ANALYSIS_PROJECTS/")
pathToThisProject            <- paste0(pathToANALYSIS_PROJECTS,thisProject)
pathToData                   <- paste0(pathToThisProject,"data/")
pathToFigures                <- paste0(pathToThisProject,"svn/analyse/script/figures/")
pathToReport                 <- paste0(pathToThisProject,"svn/analyse/report/")
pathToRNApip_RESULT          <- paste0(pathToThisProject,"RNApip_RESULT/")
pathToFeatureCounts          <- paste0(pathToThisProject,"FeatureCounts_RESULT/")
pathToDataAnnotationFile     <- paste0(pathToRhabdoid_U830_projects,"annotation_file/")

```

```{r manage_renv, eval=FALSE}

# Initialize Renv
#renv::init()

```


```{r load_library, eval=FALSE}

# Load RNAseq library
source("load_scRNAseq_library.R")

```

```{r loading_RData, eval=FALSE}

# Load saved RData
load("cellOrigin_ATRT_scRNAseq_INI254_CD45neg.RData")

#save.image(file="cellOrigin_ATRT_scRNAseq_INI254_CD45neg.RData")

```

```{r load_plot_function}

# Load external files
source("plot_R_util.R")
source("/data/users/mandrian/rhabdoid_U830_projects/annotation_file/colors.R")

```

```{r delete_all_objects}

# Delete all objects
#rm(list = ls())

```

# INI254_CD45neg scRNA-seq

## Load data and create Seurat object

```{r load_data}

pathTo_INI254_CD45neg_cellRangerCount <- "/data/users/mandrian/rhabdoid_U830_projects/ANALYSIS_PROJECTS/cellOrigin_ATRT/data/INI254/count_CD45neg/INI254_count_T1_aggr/outs/filtered_feature_bc_matrix"

# Load 10X data
INI254_CD45neg_data <- Read10X(data.dir         = pathTo_INI254_CD45neg_cellRangerCount,
                               gene.column      = 2,
                               unique.features  = TRUE
                               )

# Create Seurat object
scRNAseq_Seurat_INI254_CD45neg <- CreateSeuratObject(counts   = INI254_CD45neg_data,
                                                     project  = "INI254_CD45neg"
                                                     )
```

## Cell QC and filtering

```{r create_table_of_nbOfCell_after_filtering}

# Create data frame
cellFiltering <- data.frame(filtering = "no filtering",
                            nbOfCells = nrow(scRNAseq_Seurat_INI254_CD45neg@meta.data)
                            )

cellFiltering

#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_INI254_CD45neg_cellFiltering.pdf"), width = 12, height = 12)
ggplot(data = cellFiltering,
       mapping = aes(x = filtering,
                     y = nbOfCells
                     ),
             ) +
    geom_bar(stat = "identity",
             fill = "black"
             ) +
    ylim(0,10000) +
    xlab(label = "\nfiltering step") +
    ylab(label = "nb of cells\n") +
    geom_text(aes(label = nbOfCells), size = 15, fontface = "bold", vjust = -0.1) +
    theme_bw() +
    theme(axis.text = element_text(size = 30),
          axis.title = element_text(size = 30, face = "bold"),
          legend.position = "none"
          )
#dev.off()

```

### Number of genes and number of RNA (UMIs)

```{r number_of_genes_per_cell}

head(scRNAseq_Seurat_INI254_CD45neg@meta.data)

# Set minimum number of genes
min_nFeature <- round(quantile(scRNAseq_Seurat_INI254_CD45neg@meta.data$nFeature_RNA,
         probs = 0.625
         ))
min_nFeature

# Density distribution of total RNA count
#pdf(file = paste0(pathToFigures,"histogram_scRNAseq_INI254_CD45neg_nbRNA.pdf"), width = 12, height = 12)
ggplot(data = scRNAseq_Seurat_INI254_CD45neg@meta.data) +
    geom_histogram(mapping = aes(x = nCount_RNA),
                   fill = "yellow",
                   color = "black",
                   binwidth = 1000
                   ) +
    xlab(label = "\nnb of RNA (UMI)") +
ylab(label = "density\n") +
    theme_bw() +
    theme(axis.text = element_text(size = 30),
          axis.title = element_text(size = 30,
                                    face = "bold"
                                    )
          )
#dev.off()

# Density distribution of detected genes
#pdf(file = paste0(pathToFigures,"histogram_scRNAseq_INI254_CD45neg_nbGenes.pdf"), width = 12, height = 12)
ggplot() +
    geom_histogram(data = filter(scRNAseq_Seurat_INI254_CD45neg@meta.data, nFeature_RNA >= 500),
                   mapping = aes(x = nFeature_RNA),
                   fill = "blue",
                   color = "black",
                   binwidth = 200
                   ) +
    geom_histogram(data = filter(scRNAseq_Seurat_INI254_CD45neg@meta.data, nFeature_RNA < 500),
                   mapping = aes(x = nFeature_RNA),
                   fill = "grey",
                   color = "black",
                   binwidth = 200
                   ) +
    xlab(label = "\nnb of genes") +
ylab(label = "density\n") +
    theme_bw() +
    theme(axis.text = element_text(size = 30),
          axis.title = element_text(size = 30,
                                    face = "bold"
                                    )
          )
#dev.off()

# Scatter plot nCount_RNA and nFeature_RNA
p <- ggplot() +
    geom_point(data = scRNAseq_Seurat_INI254_CD45neg@meta.data,
               mapping = aes(x = nCount_RNA,
                             y = nFeature_RNA
                             ),
               pch = 21,
               color = "darkgreen"
               ) +
    geom_point(data = filter(.data = scRNAseq_Seurat_INI254_CD45neg@meta.data,
                             nFeature_RNA < min_nFeature
                             ),
               mapping = aes(x = nCount_RNA,
                             y = nFeature_RNA
                             ),
               pch = 21,
               color = "grey"
               ) +
    xlab(label = "\nRNA count (UMI)") +
ylab(label = "nb of genes\n") +
theme_bw() +
geom_hline(yintercept = min_nFeature,
           color = "red",
           lty = 2
           ) +
    theme(axis.text = element_text(size = 30),
          axis.title = element_text(size = 30, face = "bold")
          )
#pdf(file = paste0(pathToFigures,"scatterPlot_scRNAseq_INI254_CD45neg_nbRNA_vs_nbGene.pdf"), width = 12, height = 12)
ggMarginal(p        = p,
           type     = "density",
           margins  = "both",
           xparams  = list(fill = "yellow"),
           yparams  = list(fill = "blue")
           )
#dev.off()

# Add UMIs/gene to meta.data
nCount_per_Feature <- scRNAseq_Seurat_INI254_CD45neg@meta.data$nCount_RNA/scRNAseq_Seurat_INI254_CD45neg@meta.data$nFeature_RNA

scRNAseq_Seurat_INI254_CD45neg <- AddMetaData(object    = scRNAseq_Seurat_INI254_CD45neg,
                                              metadata  = nCount_per_Feature,
                                              col.name  = "nCount_per_Feature"
                                              )

head(scRNAseq_Seurat_INI254_CD45neg@meta.data)

# Density distribution of the number of UMIs per gene
#pdf(file = paste0(pathToFigures,"histogram_scRNAseq_INI254_CD45neg_nbRNA_per_gene.pdf"), width = 12, height = 12)
ggplot(data = scRNAseq_Seurat_INI254_CD45neg@meta.data) +
    geom_histogram(mapping = aes(x = nCount_per_Feature,
                                 y = ..density..
                                 ),
                   color = "darkgreen",
                   fill = "white"
                   ) +
        geom_density(mapping = aes(x = nCount_per_Feature),
                     color = "red",
                     size = 2
                 ) +
    xlab(label = "\nUMIs/gene") +
ylab(label = "density\n") +
    theme_bw() +
    theme(axis.text = element_text(size = 30),
          axis.title = element_text(size = 30,
                                    face = "bold"
                                    )
          )
#dev.off()

# Number of gene filtering
scRNAseq_Seurat_INI254_CD45neg_nbGenes <- subset(x = scRNAseq_Seurat_INI254_CD45neg,
                                                 subset = nFeature_RNA > min_nFeature
                                                 )

# Append cell filtering table
cellFiltering_nbGenes <- rbind(cellFiltering,
                               data.frame(filtering = "nb genes",
                                          nbOfCells = nrow(scRNAseq_Seurat_INI254_CD45neg_nbGenes@meta.data)
                                          )
                               )

cellFiltering_nbGenes

# Barplot cell filtering table
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_INI254_CD45neg_cellFiltering_nbGenes.pdf"), width = 12, height = 12)
ggplot(data = cellFiltering_nbGenes,
       mapping = aes(x = filtering,
                     y = nbOfCells,
                     fill = filtering
                     ),
             ) +
    geom_bar(stat = "identity"
             ) +
    ylim(0,10000) +
    xlab(label = "\nfiltering step") +
    ylab(label = "nb of cells\n") +
    scale_fill_manual(values = c("black","darkgreen")) +
    geom_text(aes(label = nbOfCells), size = 15, fontface = "bold", vjust = -0.1) +
    theme_bw() +
    theme(axis.text = element_text(size = 30),
          axis.title = element_text(size = 30, face = "bold"),
          legend.position = "none"
          )
#dev.off()

```

### Mitochondrial genes

```{r QC_filtering}

# Import mitochondrial gene info
pathToMitoGenes <- paste0(pathToData,"/mitochondrialRNA/group-1974.csv")

mitoGenes <- read.table(file         = pathToMitoGenes,
                        skip         = 1,
                        header       = TRUE,
                        quote        = '"',
                        sep          = ",",
                        check.names  = FALSE
                        )

head(mitoGenes)

# Mitocondrial genes percentage
scRNAseq_Seurat_INI254_CD45neg_nbGenes[["percent.mito"]] <- PercentageFeatureSet(object   = scRNAseq_Seurat_INI254_CD45neg_nbGenes,
                                                                                 features = as.vector(as.character(mitoGenes$`Approved symbol`)),
                                                                                 assay = "RNA"
                                                                                 )

head(scRNAseq_Seurat_INI254_CD45neg_nbGenes@meta.data)

# Set percentage of count falling in mitochondrial genes
max_percOfMitoGenes <- 5

# Scatter plot
p <- ggplot() +
    geom_point(data = scRNAseq_Seurat_INI254_CD45neg_nbGenes@meta.data,
               mapping = aes(x = nCount_RNA,
                             y = percent.mito
                             ),
               pch = 21,
               color = "grey"
               ) +
    geom_point(data = filter(.data = scRNAseq_Seurat_INI254_CD45neg_nbGenes@meta.data,
                             percent.mito < max_percOfMitoGenes
                             ),
               mapping = aes(x = nCount_RNA,
                             y = percent.mito
                             ),
               pch = 21,
               color = "blue"
               ) +
    xlab(label = "\nnb of RNA (UMI count)") +
ylab(label = "\n% mito. genes\n") +
geom_hline(yintercept = max_percOfMitoGenes,
           color = "red",
           lty = 2
           ) +
    theme_bw() +
    theme(axis.text = element_text(size = 20),
          axis.title = element_text(size = 20,
                                    face = "bold"
                                    )
          )
#pdf(file = paste0(pathToFigures,"scatterPlot_scRNAseq_INI254_CD45neg_nbRNA_vs_percMito.pdf"), width = 12, height = 12)
ggMarginal(p        = p,
           type     = "density",
           margins  = "y",
           fill     = "violet"
           )
#dev.off()

# Number of feature detected for each cell
#max_nFeature <- mean(scRNAseq_Seurat_INI254_CD45neg@meta.data$nFeature_RNA) + 2*sd(scRNAseq_Seurat_INI254_CD45neg@meta.data$nFeature_RNA)

# Mitochondrial gene filtering
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito <- subset(x = scRNAseq_Seurat_INI254_CD45neg_nbGenes,
                                                  subset = percent.mito < max_percOfMitoGenes
                                )

# Append cell filtering table
cellFiltering_nbGenes_percMito <- rbind(cellFiltering_nbGenes,
                                        data.frame(filtering = "perc. mito.",
                                                   nbOfCells = nrow(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito@meta.data)
                                                   )
                                        )

cellFiltering_nbGenes_percMito

# Barplot cell filtering table
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_INI254_CD45neg_cellFiltering_nbGenes_percMito.pdf"), width = 12, height = 8)
ggplot(data = cellFiltering_nbGenes_percMito,
       mapping = aes(x = reorder(filtering, -nbOfCells),
                     y = nbOfCells,
                     fill = filtering
                     ),
             ) +
    geom_bar(stat = "identity"
             ) +
    ylim(0,10000) +
    xlab(label = "\nfiltering step") +
    ylab(label = "nb of cells\n") +
    scale_fill_manual(values = c("darkgreen","black","green")) +
    geom_text(aes(label = nbOfCells), size = 15, fontface = "bold", vjust = -0.1) +
    theme_bw() +
    theme(axis.text = element_text(size = 20),
          axis.title = element_text(size = 20, face = "bold"),
          legend.position = "none"
          )
#dev.off()

```

## Data normalization

```{r data_normalization}

# Normalization
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm <- NormalizeData(object                = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito,
                                                                      normalization.method  = "LogNormalize",
                                                                      scale.factor          = 10000
                                                                      )

```

# Assign cell-cycle scores

```{r assign_cellCycle_scores}

# Cell cycle gene markers (present in Seurat)
str(cc.genes)
length(cc.genes$s.genes)
length(cc.genes$g2m.genes)

cc.genes$g2m.genes
cc.genes$s.genes

# Compute cell cycle scoring  
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm <- CellCycleScoring(object        = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm,
                                   s.features    = cc.genes$s.genes,
                                   g2m.features  = cc.genes$g2m.genes
                                   )

# Check meta data
colnames(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm@meta.data)
head(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm@meta.data)

# Plot covariate distribution
plot(density(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm@meta.data$S.Score))

plot(density(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm@meta.data$G2M.Score))

```

# Pre-analysis

This part of analysis aims to (1) find the minimun number of PCs and (2) look at the effect of covariates on clustering.


## Find min number of PCs

```{r feature_selection}

# Highly variable genes
nFeatures <- 2000
variableFeatures_method <- "vst"
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA <- FindVariableFeatures(object            = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm,
                                                                         selection.method  = variableFeatures_method,
                                                                         nfeatures         = nFeatures
                                                                         )

# Mean-variance scatter plot
#pdf(file = paste0(pathToFigures,"variableFeaturePlot_scRNAseq_INI254_CD45neg_PCA.pdf"), width = 12, height = 8)
VariableFeaturePlot(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA)
#dev.off()

nrow(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA@assays$RNA)

# Data scaling 
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA <- ScaleData(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA,
                                                              features = rownames(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA)
                                                              )

# PCA
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA <- RunPCA(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA)

# PCs elbow plot
PCmin <- 16
#pdf(file = paste0(pathToFigures,"elbowPlot_scRNAseq_INI254_CD45neg_PCA.pdf"), width = 12, height = 12)
ElbowPlot(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA,
          ndims = 50
          ) +
    geom_vline(xintercept = PCmin, lty = 2, color = "blue") +
    theme_bw() +
    theme(axis.text = element_text(size = 20),
          axis.title = element_text(size = 20, face = "bold"),
          legend.position = "none"
          )
#dev.off()

# Heatmap of each PC
#pdf(file = paste0(pathToFigures,"dimHeatmap_scRNAseq_INI254_CD45neg_PCA.pdf"), width = 15, height = 9)
DimHeatmap(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA,
           dims = 1:15,
           ncol = 5
           )
#dev.off()

# Plot PCA
#pdf(file = paste0(pathToFigures,"pca_scRNAseq_INI254_CD45neg_PCA.pdf"), width = 12, height = 12)
DimPlot(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA,
        reduction = "pca"
        ) +
    theme_bw() +
    theme(axis.text = element_text(size = 20),
          axis.title = element_text(size = 20, face = "bold"),
          legend.position = "none"
          )
#dev.off()

```

## Plot PCA and covariates

```{r PCA_and_covariates}

# Extract matrix
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix <- as.matrix(GetAssayData(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA,
                                                                                      slot = "scale.data"                                              )                                                                 )

head(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix[,1:3])

# Plot density of scaled data
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix_melt <- melt(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix)

head(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix_melt)

colnames(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix_melt) <- c("gene","cell","scaledExprs")


plot(density(exp(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix)))

# Get HVF
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_mvGenes <- HVFInfo(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA,
                                                               selection.method = "vst"
                                                               )

head(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_mvGenes)
nrow(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_mvGenes)

# Get most variable genes
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_mvGenes[order(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_mvGenes$variance.standardized, decreasing = TRUE),] %>% .[1:nFeatures,] -> mvGenes

# Subset matrix
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix_mvGenes <- scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix[as.vector(rownames(mvGenes)),]

nrow(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix_mvGenes)

# PCA analysis
PC        <- prcomp(data.frame(t(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_matrix_mvGenes)))
eigVal    <- get_eigenvalue(PC)
eigValPer <- round(eigVal$variance.percent, digits = 2)

pci       <- data.frame(PC$x,scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm@meta.data)

# Plot eigen value
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_INI254_CD45neg_PCA_eigValPer.pdf"), width = 10, height = 10)
    #mutate(PC = paste0("PC",PC)) %>%
ggplot() +
    geom_bar(data = data.frame(PC = 1:50, eigValPer = eigValPer[1:50]),
             mapping = aes(x = PC, y = eigValPer),
             stat = "identity"
             ) +
    geom_bar(data = data.frame(PC = 1:PCmin, eigValPer = eigValPer[1:PCmin]),
             mapping = aes(x = PC, y = eigValPer),
             stat = "identity",
             fill = "blue"
             ) +
    xlab(label = "\nPC") +
    ylab(label = "% variability (eigen value)\n") +
    theme_bw() +
    theme(axis.text = element_text(size = 20),
          axis.title = element_text(size = 20, face = "bold"),
          legend.position = "none"
          )
#dev.off()

# Plot PCA and nb of UMIs
#pdf(file = paste0(pathToFigures,"pca_scRNAseq_INI254_CD45neg_PCA_nbUMIs.pdf"), width = 12, height = 10)
ggplot(data = pci,
       mapping = aes(x = PC1, y = PC2)
       ) +
    geom_point(mapping = aes(color = nCount_RNA),
               size    = 1
               ) +
    scale_colour_continuous(name = "nb of UMIs",
                            type = "viridis") +
    xlab(paste0("PC1 (",eigValPer[1]," %)")) +
    ylab(paste0("PC2 (",eigValPer[2]," %)")) +
    ggtitle("PCA") +
    theme_bw()+
    theme(text              = element_text(size = 20),
          legend.title = element_text(face = "bold"),
          legend.key.height = unit(1.5,"cm")
          )
#dev.off()

# Plot PCA and nb of genes
#pdf(file = paste0(pathToFigures,"pca_scRNAseq_INI254_CD45neg_PCA_nbGenes.pdf"), width = 12, height = 10)
ggplot(data = pci,
       mapping = aes(x = PC1, y = PC2)
       ) +
    geom_point(mapping = aes(color = nFeature_RNA),
               size    = 1
               ) +
    scale_colour_continuous(name = "nb of genes",
                            type = "viridis"
                            ) +
    xlab(paste0("PC1 (",eigValPer[1]," %)")) +
    ylab(paste0("PC2 (",eigValPer[2]," %)")) +
    ggtitle("PCA") +
    theme_bw()+
    theme(text              = element_text(size = 20),
          legend.title = element_text(face = "bold"),
          legend.key.height = unit(1.5,"cm")
          )
#dev.off()

# Plot PCA and nb RNA per gene
#pdf(file = paste0(pathToFigures,"pca_scRNAseq_INI254_CD45neg_PCA_nbRNA_per_gene.pdf"), width = 12, height = 10)
ggplot(data = pci,
       mapping = aes(x = PC1, y = PC2)
       ) +
    geom_point(mapping = aes(colour = nCount_per_Feature),
               size    = 1
               ) +
    scale_colour_continuous(name = "UMIs/gene",
                            type = "viridis"
                            ) +
    xlab(paste0("PC1 (",eigValPer[1]," %)")) +
    ylab(paste0("PC2 (",eigValPer[2]," %)")) +
    ggtitle("PCA") +
    theme_bw()+
    theme(text              = element_text(size = 20),
          legend.title = element_text(face = "bold"),
          legend.key.height = unit(1.5,"cm")
          )
#dev.off()

# Plot PCA and % mito RNA
#pdf(file = paste0(pathToFigures,"pca_scRNAseq_INI254_CD45neg_PCA_percMito.pdf"), width = 12, height = 10)
ggplot(data = pci,
       mapping = aes(x = PC1, y = PC2)
       ) +
    geom_point(mapping = aes(color = percent.mito),
               size    = 1
               ) +
    scale_colour_continuous(name = "% mito",
                            type = "viridis"
                            ) +
    xlab(paste0("PC1 (",eigValPer[1]," %)")) +
    ylab(paste0("PC2 (",eigValPer[2]," %)")) +
    ggtitle("PCA") +
    theme_bw()+
    theme(text              = element_text(size = 20),
          legend.title = element_text(face = "bold"),
          legend.key.height = unit(1.5,"cm")
          )
#dev.off()

# Plot PCA and cell cycle phase
#pdf(file = paste0(pathToFigures,"pca_scRNAseq_INI254_CD45neg_PCA_cellCyclePhase.pdf"), width = 12, height = 10)
ggplot(data = pci,
       mapping = aes(x = PC1, y = PC2)
       ) +
    geom_point(mapping = aes(color = Phase),
               size    = 1#,
               #pch     = 21
               ) +
    #scale_fill_continuous(type = "viridis") +
    xlab(paste0("PC1 (",eigValPer[1]," %)")) +
    ylab(paste0("PC2 (",eigValPer[2]," %)")) +
    ggtitle("PCA") +
    theme_bw()+
    theme(text              = element_text(size = 30),
          legend.title = element_text(face = "bold"),
          legend.key.height = unit(1.5,"cm")
          )
#dev.off()

```

# Clustering (Using PCmin as identified above)

```{r clustering}

# Construct KNN graph
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust <- FindNeighbors(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA,
                                                                                dims = 1:PCmin
                                                                                )

# Clustering by modularity optimization technique
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust <- FindClusters(object     = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
                                                                       resolution = seq(from  = 0,
                                                                                        to    = 2,
                                                                                        by    = 0.1
                                                                                        )
                                                                       )      

# Save seurat object
saveRDS(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
        file   = paste0(pathToReport,"scRNAseq_Seurat_INI254_CD45neg.rds"
                        )
        )

# Save as loom object
as.loom(x         = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
        filename  = "cellOrigin_ATRT_scRNAseq_INI254_CD45neg.loom"
        )
    
```

## Choose the best resolution of clustering

```{r choose_resolution}

# Choose resolution
#pdf(file = paste0(pathToFigures,"clustree_scRNAseq_INI254_CD45neg.pdf"), width = 10, height = 12)
clustree(x               = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust@meta.data,
         prefix          = "RNA_snn_res.",
         #node_colour     = "sc3_stability",
         layout          = "sugiyama",
         use_core_edges  = TRUE,
         color           = "red"
         )
#dev.off()

choosenResolution <- 0.4

# Clustering by modularity optimization technique
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution <- FindClusters(object     = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
                                                                                                 resolution = choosenResolution
                                                                                                 )      

# Number of cell per cluster
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_Seurat_INI254_CD45neg_numberOfcellsPerCluster.pdf"), width = 16, height = 10)
ggplot(data = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution@meta.data
       ) +
    stat_count(mapping = aes(x     = RNA_snn_res.0.2,
                             fill  = Phase
                             )
               ) +
    geom_text(mapping  = aes(x     = RNA_snn_res.0.2,
                             label = stat(count)
                             ),
              stat     = "count",
              vjust    = -0.25,
              size     = 7.5,
              face     = "bold"
              ) +
    xlab("\nCluster") +
    ylab("# of cells\n") +
    theme_bw() +
        theme(text = element_text(size = 40),
              legend.position = c(0.85,0.95),
              legend.direction = "horizontal"
              )
#dev.off()

# Export meta data (for SCENIC analysis)
head(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust@meta.data)

write.table(x      = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust@meta.data,
            file   = "scRNAseq_Seurat_INI254_CD45neg_PCmin_metaData.txt",
            sep    = ",",
            quote  = FALSE
            )

```

## UMAP

```{r UMAP}

# Run UMAP
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust <- RunUMAP(object       = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
                                                                          dims         = 1:PCmin,
                                                                          n.neighbors  = 30
                                                                          )

# Plot UMAP
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_INI254_CD45neg_clust.pdf"), width = 12, height = 12)
DimPlot(object      = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
        reduction   = "umap",
        pt.size     = 3,
        group.by    = paste0("RNA_snn_res.",choosenResolution),
        label       = TRUE,
        label.size  = 15
        ) +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20)
          )
#dev.off()

```

## UMAP and covariates

```{r umap_and_covariates}

# UMAP and nb of UMIs
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_INI254_CD45neg_clust_nbUMIs.pdf"), width = 12, height = 12)
FeaturePlot(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
            features = "nFeature_RNA",
            reduction = "umap",
            pt.size = 3
            ) +
    ggtitle(label = NULL) +
    theme_bw() +
    theme(legend.position = c(0.85,0.85),
          text = element_text(size = 20)
          )
#dev.off()

# UMAP and nb of genes
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_INI254_CD45neg_clust_nbGenes.pdf"), width = 12, height = 12)
FeaturePlot(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
            features = "nCount_RNA",
            reduction = "umap",
            pt.size = 3
            ) +
    ggtitle(label = NULL) +
    theme_bw() +
    theme(legend.position = c(0.85,0.85),
          text = element_text(size = 20)
          )
#dev.off()

# UMAP and nb of UMIs per gene
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_INI254_CD45neg_clust_nbRNA_per_gene.pdf"), width = 12, height = 12)
FeaturePlot(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
            features = "nCount_per_Feature",
            reduction = "umap",
            pt.size = 3
            ) +
    ggtitle(label = NULL) +
    theme_bw() +
    theme(legend.position = c(0.85,0.85),
          text = element_text(size = 20)
          )
#dev.off()

# UMAP and percentage of mitochondrial genes
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_INI254_CD45neg_clust_percMito.pdf"), width = 12, height = 12)
FeaturePlot(object = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
            features = "percent.mito",
            reduction = "umap",
            pt.size = 3
            ) +
    ggtitle(label = NULL) +
    theme_bw() +
    theme(legend.position = c(0.85,0.85),
          text = element_text(size = 20)
          )
#dev.off()

# UMAP and cell cycle phase
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_INI254_CD45neg_clust_cellCyclePhase.pdf"), width = 12, height = 12)
DimPlot(object      = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
        reduction   = "umap",
        pt.size     = 3,
        group.by    = "Phase",
        label       = TRUE,
        label.size  = 15
        ) +
    ggtitle(label = NULL) +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20)
          )
#dev.off()

```

## Boxplot of covariates per cluster

```{r boxplot_and_covariates}

# Boxplot and nb of UMIs
#pdf(file = paste0(pathToFigures,"boxplot_scRNAseq_INI254_CD45neg_clust_nbUMIs.pdf"), width = 12, height = 12)
 ggplot(data = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust@meta.data) +
    geom_jitter(mapping = aes(x      = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                              y      = log10(nCount_RNA),
                              color  = eval(as.name(paste0("RNA_snn_res.",choosenResolution)))
                              ),
                size    = 1
                ) +
    stat_summary(mapping = aes(x = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                               y = log10(nCount_RNA)
                              ),
                 fun     = median,
                 geom    = "crossbar",
                 width   = 0.75,
                 color   = "blue"
                 ) +
    xlab(label = "\nCluster") +
    ylab(label = "log10(UMI count)\n")+
    ylim(2,5) +
    theme_bw() +
    theme(legend.position = "none",
          text            = element_text(size = 40)
          )
#dev.off()

# Boxplot and nb of genes
#pdf(file = paste0(pathToFigures,"boxplot_scRNAseq_INI254_CD45neg_clust_nbGenes.pdf"), width = 12, height = 12)
ggplot(data = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust@meta.data) +
    geom_jitter(mapping = aes(x      = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                              y      = nFeature_RNA,
                              color  = eval(as.name(paste0("RNA_snn_res.",choosenResolution)))
                              ),
                size    = 1
                ) +
    stat_summary(mapping = aes(x = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                               y = nFeature_RNA
                              ),
                 fun     = median,
                 geom    = "crossbar",
                 width   = 0.75,
                 color   = "blue"
                 ) +
    xlab(label = "\nCluster") +
    ylab(label = "# genes\n")+
    #ylim(2,5) +
    theme_bw() +
    theme(legend.position = "none",
          text            = element_text(size = 40)
          )
#dev.off()

# Boxplot and nb of UMIs per gene
#pdf(file = paste0(pathToFigures,"boxplot_scRNAseq_INI254_CD45neg_clust_nbRNA_per_gene.pdf"), width = 12, height = 12)
ggplot(data = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust@meta.data) +
    geom_jitter(mapping = aes(x      = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                              y      = nCount_per_Feature,
                              color  = eval(as.name(paste0("RNA_snn_res.",choosenResolution)))
                              ),
                size    = 1
                ) +
    stat_summary(mapping = aes(x = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                               y = nCount_per_Feature
                              ),
                 fun     = median,
                 geom    = "crossbar",
                 width   = 0.75,
                 color   = "blue"
                 ) +
    xlab(label = "\nCluster") +
    ylab(label = "UMIs per gene\n")+
    #ylim(2,5) +
    theme_bw() +
    theme(legend.position = "none",
          text            = element_text(size = 40)
          )
#dev.off()

# Boxplot and percentage of mitochondrial genes
#pdf(file = paste0(pathToFigures,"boxplot_scRNAseq_INI254_CD45neg_clust_percMito.pdf"), width = 12, height = 12)
ggplot(data = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust@meta.data) +
    geom_jitter(mapping = aes(x      = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                              y      = percent.mito,
                              color  = eval(as.name(paste0("RNA_snn_res.",choosenResolution)))
                              ),
                size    = 1
                ) +
    stat_summary(mapping = aes(x = eval(as.name(paste0("RNA_snn_res.",choosenResolution))),
                               y = percent.mito
                               ),
                 fun     = median,
                 geom    = "crossbar",
                 width   = 0.75,
                 color   = "blue"
                 ) +
    xlab(label = "\nCluster") +
    ylab(label = "% mito genes\n")+
    #ylim(2,5) +
    theme_bw() +
    theme(legend.position = "none",
          text            = element_text(size = 40)
          )
#dev.off()

```


## Differential analysis

### One versus all (finding cluster biomarkers)

```{r diffAnalysis}

# Set thresholds
min_pct          <- 0.25
min_logfc        <- 0.50
min_diffPct      <- -Inf
min_cellsFeature <- 3

# Find markers for each cluster (versus all others)
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers <- FindAllMarkers(object            = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution,
                                                                                 min.pct           = min_pct,
                                                                                 logfc.threshold   = min_logfc,
                                                                                 min.diff.pct      = min_diffPct,
                                                                                 only.pos          = TRUE,
                                                                                 min.cells.feature = min_cellsFeature
                                                                                 )

head(scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers)

# Plot the number of markers detected for each cluster
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_INI254_CD45neg_numberOfMarkers.pdf"), width = 16, height = 10)
ggplot(data = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers,
       mapping = aes(x = cluster,
                     fill = cluster
                     )
       ) +
    stat_count() +
    geom_text(mapping = aes(label = stat(count)),
              stat = "count",
              vjust = -0.25,
              size = 10#,
              #fontface = "bold"
              ) +
    xlab("\nCluster") +
    ylab("# of markers\n") +
    theme_bw() +
        theme(text = element_text(size = 40),
              legend.position = "none"
              )
#dev.off()

# Look at the top n markers for each cluster
nTop <- 5
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers %>% group_by(cluster) %>% top_n(n = nTop, wt = avg_logFC)

# Select top markers for each cluster
nTop_markers <- scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers %>% group_by(cluster) %>% top_n(n = nTop, wt = avg_logFC)

# Heatmap of all markers
#pdf(file = paste0(pathToFigures,"heatmap_scRNAseq_INI254_CD45neg_clusterMarkers.pdf"), width = 16, height = 10)
DoHeatmap(object   = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution,
          features = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers$gene,
          assay    = "RNA"
          ) +
    NoLegend() +
    theme(axis.text.y = element_blank())
#dev.off()

# Heatmap of top markers
#pdf(file = paste0(pathToFigures,"heatmap_scRNAseq_INI254_CD45neg_clusterTopMarkers.pdf"), width = 16, height = 12)
DoHeatmap(object   = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution,
          features = nTop_markers$gene,
          assay    = "RNA"
          ) +
    NoLegend() +
    theme(axis.text.y = element_text(size=20))
#dev.off()

# Plot top markers for each cluster
#pdf(file = paste0(pathToFigures,"volcanoPlot_scRNAseq_INI254_CD45neg_clusterTopMarkers.pdf"), width = 16, height = 10)
nTop = 20
scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers %>%
    group_by(cluster) %>%
    top_n(n = 20, wt = avg_logFC) %>%
    ggplot() +
    geom_point(mapping = aes(x = avg_logFC,
                             y = -log10(p_val_adj),
                             size = pct.1 - pct.2 
                             ),
               pch = 21,
               fill = "blue",
               alpha = 0.5
               ) +
    facet_wrap(facets = ~ cluster, nrow = 2) +
    geom_text_repel(mapping = aes(x = avg_logFC,
                                  y = -log10(p_val_adj),
                                  label = gene
                                  ),
                    size = 3
                    ) +
    xlab("\navg_logFC") +
    ylab("-log10(p_val_adj)\n") +
    theme_bw() +
    theme(text = element_text(size = 20),
          strip.background = element_rect(fill = "green"),
          strip.text = element_text(face = "bold")
          )
#dev.off()

# Export marker table
write.table(x           = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution.markers,
            file        = paste0(pathToReport,"scRNAseq_INI254_CD45neg_cluster_geneMarkers.txt"),
            sep         = "\t",
            quote       = FALSE
            )

```

### SMARCB1 expression

```{r SMARCB1_expression}

# SMARCB1 expression in UMAP
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_INI254_CD45neg_SMARCB1.pdf"), width = 12, height = 12)
FeaturePlot(object   = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust,
            features = "SMARCB1",
            pt.size  = 3
            ) +
    ggtitle(label = "SMARCB1") +
    theme_bw() +
    theme(legend.position = c(0.85,0.85),
          text = element_text(size = 20),
          plot.title = element_text(hjust = 0.5, face = "bold")
          )
#dev.off()

# SMARCB1 expression in violin plot
#pdf(file = paste0(pathToFigures,"violinPlot_scRNAseq_INI254_CD45neg_SMARCB1.pdf"), width = 12, height = 12)
VlnPlot(object   = scRNAseq_Seurat_INI254_CD45neg_nbGenes_percMito_norm_PCA_clust_choosenResolution,
        features = "SMARCB1",
        pt.size = 0
        ) +
    ggtitle(label = "SMARCB1") +
    xlab("\nCluster") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20),
          plot.title = element_text(hjust = 0.5, face = "bold")
          )
#dev.off()

```

# Session info

```{r sessionInfo}

# Session info
sessionInfo()

```
