---
title: "Human ATRT RNAseq data"
author: "Mamy Andrianteranagna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 9
    keep_md: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
    number_sections: true
    code_folding: hide
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, include=TRUE, cache=FALSE, error=TRUE)

```

```{r paths}

thisProject                  <- "cellOrigin_ATRT/"

# Paths
pathToRhabdoid_U830_projects <- "/bioinfo/users/mandrian/rhabdoid_U830_projects/"
pathToProgrammes             <- paste0(pathToRhabdoid_U830_projects,"programmes/")
#pathToRpackages              <- paste0(pathToProgrammes,"Rpackages/")
#pathToRpackages              <- paste0(pathToProgrammes,"Rpackages_R.3.6/")
#pathToRpackages              <- paste0(pathToProgrammes,"Rpackages_R.3.5/")
pathToANALYSIS_PROJECTS      <- paste0(pathToRhabdoid_U830_projects,"ANALYSIS_PROJECTS/")
pathToThisProject            <- paste0(pathToANALYSIS_PROJECTS,thisProject)
pathToData                   <- paste0(pathToThisProject,"data/")
pathToFigures                <- paste0(pathToThisProject,"svn/analyse/script/figures/")
pathToReport                 <- paste0(pathToThisProject,"svn/analyse/report/")
pathToRNApip_RESULT          <- paste0(pathToThisProject,"RNApip_RESULT/")
pathToFeatureCounts          <- paste0(pathToThisProject,"FeatureCounts_RESULT/")

pathToRNApipResults          <-  paste0(pathToThisProject,"RNApip_RESULT/")

pathToDataAnnotationFile     <- paste0(pathToRhabdoid_U830_projects,"annotation_file/")

```

```{r manage_renv, eval=FALSE}

# Initialize Renv
#renv::init()

```


```{r load_library, eval=FALSE}

# Load RNAseq library
source("load_scRNAseq_library.R")

```

```{r loading_RData, eval=FALSE}

# Load saved RData
load("cellOrigin_ATRT_scRNAseq_combinedSamples.RData")

#save.image(file="cellOrigin_ATRT_scRNAseq_combinedSamples.RData")

```

```{r load_plot_function}

# Load external files
source("plot_R_util.R")
source("/data/users/mandrian/rhabdoid_U830_projects/annotation_file/colors.R")

```

```{r delete_all_objects}

# Delete all objects
rm(list = ls())

```

```{r paths}

# def des path
main_dir = "/data/kdi_prod/project_result/1002/02.00/svn/analyse/results/human_sample_merge"

```

```{r params}

# def des paramètres
select.dimension = 50
nFeatures        = 2000

```

# Import and process individual sample seurat object

```{r import_singleSample_seuratObject}

Patient_IDs <- c("INI254","INI255","INI267")

sampleSeuratObj_list <- list()
for(sample in Patient_IDs)
{
    sampleSeuratObj = readRDS(file = file.path(main_dir, paste(sample, "Norm_features-mito.data.rds", sep = "_")))
    
    # Rename the cells
    sampleSeuratObj <- RenameCells(object = sampleSeuratObj,
                                   add.cell.id = sample
                                   )

                                        # Normalization (Log Normalize method)
    sampleSeuratObj <- NormalizeData(object                = sampleSeuratObj,
                                     normalization.method  = "LogNormalize",
                                     scale.factor          = 1e4
                                     )

    # Variable features
    sampleSeuratObj <- FindVariableFeatures(object = sampleSeuratObj, 
                                            selection.method = 'vst', 
                                            mean.cutoff = c(-Inf, Inf),
                                            dispersion.cutoff = c(0.5, Inf),
                                            nfeatures = nFeatures,
                                            verbose = TRUE)

    sampleSeuratObj_list[[length(sampleSeuratObj_list)+1]] <- sampleSeuratObj
}
names(sampleSeuratObj_list) <- Patient_IDs

```

# Data integration 

```{r data_integration}

# Find anchors (CCA)
anchors <- FindIntegrationAnchors(object.list      = sampleSeuratObj_list,
                                  anchor.features  = nFeatures,
                                  dims             = 1:select.dimension
                                  )
anchors

# Data integration 
integratedData <- IntegrateData(anchorset  = anchors,
                                dims       = 1:select.dimension
                                )

table(integratedData$sampleCond)
table(integratedData$orig.ident)

# Barplot of number of cell per sample
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA_numberOfCells.pdf"), width = 12, height = 12)
table(integratedData$orig.ident) %>%
    data.frame() %>%
    ggplot(mapping = aes(x = Var1,
                         y = Freq,
                         fill = Var1
                         )
             ) +
    geom_bar(stat = "identity") +
    geom_text(mapping = aes(label = Freq),
              size = 10,
              fontface = "bold",
              vjust = -0.25
              ) +
    xlab("\nTumor sample") +
    ylab("# of cells\n") +
theme_bw() +
    theme(text = element_text(size = 30, face = "bold"),
          legend.position = "none"
          )
#dev.off()

```

# Data scaling

```{r data_scaling}

# Scale data 
all.genes <- rownames(integratedData@assays$RNA)
integratedData <- ScaleData(object = integratedData,
                            features = all.genes
                            )
```

# Linear dimensional reduction

```{r linear_dimensional_reduction}

# Compute PCA
integratedData <- RunPCA(integratedData,
                         npcs = select.dimension
                         )

# Plot PCA
#pdf(file = paste0(pathToFigures,"pca_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA.pdf"), width = 14, height = 10)
DimPlot(object     = integratedData,
        reduction  = "pca",
        pt.size    = 2
        ) +
    theme_bw() +
    theme(text = element_text(size = 40)
          )
#dev.off()

# Plot heatmap of genes in PC1
#pdf(file = paste0(pathToFigures,"dimHeatmap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA.pdf"), width = 16, height = 10)
DimHeatmap(object    = integratedData,
           dims      = 1:10,
           cells     = 500,
           balanced  = TRUE,
           ncol      = 5
           )
#dev.off()

# Determine "singnificant" components by permutation 
integratedData <- JackStraw(object         = integratedData,
                            reduction      = "pca",
                            dims           = select.dimension,
                            num.replicate  = 100,
                            prop.freq      = 0.01
                            )

integratedData <- ScoreJackStraw(object = integratedData,
                                 dims   = 1:select.dimension
                                 )

#pdf(file = paste0(pathToFigures,"jakStrawPlot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA.pdf"), width = 12, height = 12)
JackStrawPlot(object  = integratedData, dims = 1:select.dimension)
#dev.off()

# Determine "significant" components using heuristic method
#pdf(file = paste0(pathToFigures,"elbowPlot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA.pdf"), width = 12, height = 12)
ElbowPlot(object     = integratedData,
          ndims      = select.dimension,
          reduction  = "pca"
          ) +
    theme_bw() +
    theme(text = element_text(size = 40)
          )
#dev.off()

choosenDim <- 1:10

```

# Cell clustering 

## Clustering

```{r clustering}

# Construct KNN graph 
integratedData <- FindNeighbors(object  = integratedData,
                                dims    = choosenDim,
                                assay   = "integrated"
                                )

# Clustering by modularity optimization technique
integratedData <- FindClusters(object     = integratedData,
                               resolution = seq(from  = 0,
                                                to    = 2,
                                                by    = 0.1
                                                )
                               )

```

## Choose the best resolution of clustering

```{r choose_resolution}

# Choose resolution
#pdf(file = paste0(pathToFigures,"clustree_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA.pdf"), width = 10, height = 12)
clustree(x               = integratedData@meta.data,
         prefix          = "integrated_snn_res.",
         node_colour     = "sc3_stability",
         layout          = "sugiyama",
         use_core_edges  = TRUE
         )
#dev.off()

choosenResolution <- 0.9

```

## Cluster visualization

```{r clusterVisualization}

# Run UMAP
integratedData <- RunUMAP(object       = integratedData,
                          dims         = choosenDim,
                          n.neighbors  = 30
                          )

# Plot UMAP
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA.pdf"), width = 12, height = 12)
DimPlot(object      = integratedData,
        reduction   = "umap",
        pt.size     = 2,
        group.by    = paste0("integrated_snn_res.",choosenResolution),
        label       = TRUE,
        label.size  = 10
        ) +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20)
          )
#dev.off()

```

## Extract cluster average expression of scaled matrix

```{r extract_average_matrix}

# Find cluster for choosen resolution
integratedData_choosenDim <- FindClusters(object = integratedData,
                                          resolution = choosenResolution
                                          )

# Extract matrix
matrix_integratedData_choosenDim <- AverageExpression(object = integratedData_choosenDim,
                                                      assays = "RNA",
                                                      features.assay = "RNA",
                                                      features = all.genes,
                                                      )$RNA

head(matrix_integratedData_choosenDim[,1:4])
nrow(matrix_integratedData_choosenDim)

# Export matrix
write.table(x      = matrix_integratedData_choosenDim,
            file   = paste0(pathToReport,"matrix_integratedData_ATRT_SHH_anteriorVermis.txt"),
            sep    = "\t",
            quote  = FALSE
            )

```

## Number of cell per cluster

```{r numberOfCellPerCluster}

#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA_numberOfcellsPerCluster.pdf"), width = 16, height = 10)
ggplot(data = integratedData_choosenDim@meta.data
       ) +
    stat_count(mapping = aes(x     = integrated_snn_res.0.9,
                             fill  = orig.ident
                             )
               ) +
    geom_text(mapping  = aes(x     = integrated_snn_res.0.9,
                             label = stat(count)
                             ),
              stat     = "count",
              vjust    = -0.25,
              size     = 7.5,
              face     = "bold"
              ) +
    xlab("\nCluster") +
    ylab("# of cells\n") +
    theme_bw() +
        theme(text = element_text(size = 40),
              legend.position = c(0.85,0.85)
              )
#dev.off()

```

# Differential analysis

## One versus all (finding cluster biomarkers)

```{r diffAnalysis}

DefaultAssay(integratedData_choosenDim) <- "RNA"

# Set thresholds
min_pct          <- 0.25
min_logfc        <- 0.5
min_diffPct      <- -Inf
min_cellsFeature <- 3

# Find markers for each cluster (versus all others)
integratedData_choosenDim.markers <- FindAllMarkers(object            = integratedData_choosenDim,
                                                    min.pct           = min_pct,
                                                    logfc.threshold   = min_logfc,
                                                    min.diff.pct      = min_diffPct,
                                                    only.pos          = TRUE,
                                                    min.cells.feature = min_cellsFeature
                                                    )

head(integratedData_choosenDim.markers)
nrow(integratedData_choosenDim.markers)

# Plot the number of markers detected for each cluster
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA_numberOfMarkers.pdf"), width = 16, height = 10)
ggplot(data = integratedData_choosenDim.markers,
       mapping = aes(x = cluster,
                     fill = cluster
                     )
       ) +
    stat_count() +
    geom_text(mapping = aes(label = stat(count)),
              stat = "count",
              vjust = -0.25,
              size = 10,
              face = "bold"
              ) +
    xlab("\nCluster") +
    ylab("# of markers\n") +
    theme_bw() +
        theme(text = element_text(size = 40),
              legend.position = "none"
              )
#dev.off()


# Export marker table
write.table(x           = integratedData_choosenDim.markers,
            file        = paste0(pathToReport,"scRNAseq_ATRT_SHH_anteriorVermis_cluster_geneMarkers.txt"),
            sep         = "\t",
            quote       = FALSE
            )

# Heatmap of all markers
#pdf(file = paste0(pathToFigures,"heatmap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA_clusterMarkers.pdf"), width = 16, height = 10)
DoHeatmap(object   = integratedData_choosenDim,
          features = integratedData_choosenDim.markers$gene,
          assay    = "integrated"
          ) +
    NoLegend() +
    theme(axis.text.y = element_blank())
#dev.off()

# Look at the top n markers for each cluster
nTop <- 5
integratedData_choosenDim.topMarkers <- integratedData_choosenDim.markers %>%
    group_by(cluster) %>%
    top_n(n = nTop, wt = avg_logFC)
integratedData_choosenDim.topMarkers

# Heatmap of top markers
#pdf(file = paste0(pathToFigures,"heatmap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_CCA_clusterTopMarkers.pdf"), width = 16, height = 12)
DoHeatmap(object   = integratedData_choosenDim,
          features = integratedData_choosenDim.topMarkers$gene,
          assay    = "integrated"
          ) +
    NoLegend() +
    theme(axis.text.y = element_text(size=20))
#dev.off()

```

# Cell type identification

## Tumor cells identification

### Import and explore cell types database

```{r importCellTypeMarkers}

# Import cannonical and published cell type markers
cellType_markers <- read.table(file    = "/data/users/mandrian/rhabdoid_U830_projects/annotation_file/cellType_markers.csv",
                               header  = TRUE,
                               sep     = "\t"
                               )
head(cellType_markers)

# Gene markers per cell type
group_by(.data = filter(cellType_markers,
                        Cell.type == "Microglia"
                        )
         #,Cell.type
         ,Source
         ) %>%
    summarise(number = n()) %>%
    print(n=nrow(.), width = Inf)

```

### Import and explore geneSet database

```{r importCellTypeMarkers}

# Import genSet
geneSet <- read.table(file    = "/data/users/mandrian/rhabdoid_U830_projects/geneSet.csv",
                               header  = TRUE,
                               sep     = ";"
                               )
head(geneSet)

# Gene markers per cell type
group_by(.data = geneSet,
         #,Cell.type
         ,Group
         ) %>%
    summarise(number = n()) %>%
    print(n=nrow(.), width = Inf)

```

### ATRT SHH signature expression

```{r SHH_markers_expression}

# Select SHH gene signatures
geneSet_ATRT_SHH <- filter(.data = geneSet,
                           Group == "ATRT_SHH(1)"
                           )

nrow(geneSet_ATRT_SHH)

as.vector(geneSet_ATRT_SHH$Gene)

# ATRT SHH signature gene expression in UMAP
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_ATRT_SHH_genes.pdf"), width = 14, height = 10)
FeaturePlot(object   = integratedData_choosenDim,
            features = as.vector(geneSet_ATRT_SHH$Gene),
            pt.size  = 0.25
            ) #+
    #ggtitle(label = "SMARCB1") +
    #theme_bw() +
    #theme(legend.position = "none",
          #text = element_text(size = 20),
          #plot.title = element_text(hjust = 0.5, face = "bold")
          #)
#dev.off()

# ATRT SHH signature gene expression in violin plot
#pdf(file = paste0(pathToFigures,"violinPlot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_ATRT_SHH_genes.pdf"), width = 16, height = 8)
VlnPlot(object   = integratedData_choosenDim,
        features = as.vector(geneSet_ATRT_SHH$Gene)
        ) #+
    #ggtitle(label = "SMARCB1") +
    #xlab("\nCluster") +
    #theme_bw() +
    #theme(legend.position = "none",
          #text = element_text(size = 20),
          #plot.title = element_text(hjust = 0.5, face = "bold")
          #)
#dev.off()


```

### SMARCB1 expression

```{r SMARCB1_expression}

# SMARCB1 expression in UMAP
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_SMARCB1.pdf"), width = 12, height = 12)
FeaturePlot(object   = integratedData_choosenDim,
            features = "SMARCB1",
            pt.size  = 2
            ) +
    ggtitle(label = "SMARCB1") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20),
          plot.title = element_text(hjust = 0.5, face = "bold")
          )
#dev.off()

# SMARCB1 expression in violin plot
#pdf(file = paste0(pathToFigures,"violinPlot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_SMARCB1.pdf"), width = 16, height = 8)
VlnPlot(object   = integratedData_choosenDim,
        features = "SMARCB1"#,
        #pt.size  = 2
        ) +
    ggtitle(label = "SMARCB1") +
    xlab("\nCluster") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20),
          plot.title = element_text(hjust = 0.5, face = "bold")
          )
#dev.off()

```

### Canonical markers expression

#### Immune cells canonical markers

```{r immuneCells}

# Select immune cells canonical markers
cellType_markers_immuneCell_canonical <- filter(.data = cellType_markers,
                                                Source == "canonical marker",
                                                Group == "immune cells",
                                                Organism == "human"
                                                )


as.vector(cellType_markers_immuneCell_canonical$Gene)


#!!!!!!!!!!!!!!!!!!!!!
max(as.matrix(log2(matrix_integratedData_choosenDim+1)))

# Subset average matrix of immune cell maker genes
matrix_integratedData_choosenDim_immuneCellMarkers <- matrix_integratedData_choosenDim[rownames(matrix_integratedData_choosenDim) %in% as.vector(cellType_markers_immuneCell_canonical$Gene),]

head(matrix_integratedData_choosenDim_immuneCellMarkers)

col_fun <- circlize::colorRamp2(breaks = c(0,0.5,10), colors = c("white","grey","black"))

# Circle heatmap of average expression
pdf(file = paste0(pathToFigures,"circleHeatmap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_immuneCells_canonicalMarkers.pdf"), width = 16, height = 10)
Heatmap(matrix = matrix_integratedData_choosenDim_immuneCellMarkers,
        col = col_fun,
        rect_gp = gpar(type="none"),
        cluster_columns = FALSE,
        cluster_rows = FALSE,
        show_row_names = TRUE,
        cell_fun = function(j, i, x, y, width, height, fill)
        {
            grid.rect(x = x, y = y, width = width, height = height, gp = gpar(col = "grey", fill = NA))
            grid.circle(x = x, y = y, r = (matrix_integratedData_choosenDim_immuneCellMarkers[i,j]+1)/10*min(unit.c(width, height)),
                        gp = gpar(fill = col_fun(matrix_integratedData_choosenDim_immuneCellMarkers[i,j]), col = "black")
                        )
        }
        )
dev.off()

# Heatmap of average expression
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_immuneCells_canonicalMarkers.pdf"), width = 16, height = 10)
FeaturePlot(object   = integratedData_choosenDim,
            features = unique(as.vector(cellType_markers_immuneCell_canonical$Gene)),
            pt.size  = 0.25
            ) #+
    #ggtitle(label = "SMARCB1") +
    #theme_bw() +
    #theme(legend.position = "none",
          #text = element_text(size = 20),
          #plot.title = element_text(hjust = 0.5, face = "bold")
          #)
#dev.off()

```

# Create human-mouse gene name conversion

```{r human_mouse_conversion}

# Load human and mouse ensembl datasets annotation
human <- useMart(biomart="ensembl", dataset="hsapiens_gene_ensembl")
mouse <- useMart(biomart="ensembl", dataset="mmusculus_gene_ensembl")

# Create human gene symbol and mouse orthologous gene symbol corresponding table
symbols_conversion <- getLDS(attributes  = c("hgnc_symbol"),
                             mart        = human,
                             attributesL = "mgi_symbol",
                             martL       = mouse,
                             uniqueRows  = TRUE
                             )
head(symbols_conversion)

# Add mouse orthologous gene symbol
integratedData_choosenDim.markers_plus <- merge(x      = integratedData_choosenDim.markers,
                                                y      = symbols_conversion,
                                                by.x   = "gene",
                                                by.y   = "HGNC.symbol",
                                                all.x  = TRUE,
                                                all.y  = FALSE
                                                )
head(integratedData_choosenDim.markers_plus)

```


#### Microglia gene set from Jessa2019

```{r immuneCells}

head(cellType_markers)

# Select immune cells canonical markers
cellType_markers_Microglia_Jessa2019 <- filter(.data      = cellType_markers,
                                               Source     == "Jessa et al., 2019",
                                               Cell.type  == "Microglia",
                                               Organism   == "murine"
                                               )



head(cellType_markers_Microglia_Jessa2019$Gene)

levels(integratedData_choosenDim.markers$cluster)

cluster <- vector()
k <- vector()
K <- vector()
pVal <- vector()
for (clust in levels(integratedData_choosenDim.markers$cluster))
{
    g <- filter(.data    = integratedData_choosenDim.markers_plus,
                cluster  == clust
                )$MGI.symbol
    K <- c(K,length(g))
    G <- as.vector(cellType_markers_Microglia_Jessa2019$Gene)
    # Intersect cluster gene markers and cell type markers
    gG <- intersect(g,G)
    k <- c(k,length(gG))
    # Hypergeometric test
    p <- computeHypergeometrix_pval(x = g,
                                    y = G,
                                    N = nrow(integratedData_choosenDim)
                                    )
    pVal <- c(pVal,p)
    cluster <- c(cluster,clust)
}
overlap_integratedData_markers_Microglia_Jessa2019 <- data.frame(cluster,k,K,pVal)

# Compute q values
overlap_integratedData_markers_Microglia_Jessa2019$qVal <- qvalue(p = overlap_integratedData_markers_Microglia_Jessa2019$pVal)$qvalues
overlap_integratedData_markers_Microglia_Jessa2019

# Keep cluster order for ggplot
overlap_integratedData_markers_Microglia_Jessa2019$cluster <- factor(overlap_integratedData_markers_Microglia_Jessa2019$cluster,
                                                                     levels = overlap_integratedData_markers_Microglia_Jessa2019$cluster
                                                                     )

# Plot barplot
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_markers_Microglia_Jessa2019.pdf"), width = 16, height = 8)
ggplot(data    = overlap_integratedData_markers_Microglia_Jessa2019,
       mapping = aes(x = cluster,
                     y = -log10(qVal)
                     )
       ) +
    geom_bar(stat = "identity",
             fill = "brown"
             ) +
    #coord_flip() +
    xlab("\nCluster") +
    ylab("-log10(q-value)\n") +
    #ylim(0,xlim) +
    #geom_text(mapping = aes(label=paste0("(",k,"/",K,")")),
     #         size = 10,
      #        hjust = -0.1
       #       ) +
    theme_bw() +
    theme(text         = element_text(size=40, color = "black"),
          panel.border = element_blank(),
          axis.text.x  = element_text(color = "black")
          )
#dev.off()

```

### Microglia marker Soussa2017

```{r SMARCB1_expression}

# Select immune cells canonical markers
cellType_markers_Microglia_Soussa2017 <- filter(.data = cellType_markers,
                                                Source == "Sousa et al., 2017",
                                                Cell.type == "Microglia",
                                                Organism == "murine"
                                                )



# SMARCB1 expression in UMAP
#pdf(file = paste0(pathToFigures,"umap_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_Microglia_Soussa2017.pdf"), width = 12, height = 12)
FeaturePlot(object   = integratedData_choosenDim,
            features = symbols_conversion[symbols_conversion$MGI.symbol == cellType_markers_Microglia_Soussa2017$Gene,"HGNC.symbol"],
            pt.size  = 2
            ) +
    #ggtitle(label = "") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20),
          plot.title = element_text(hjust = 0.5, face = "bold")
          )
#dev.off()

# SMARCB1 expression in violin plot
#pdf(file = paste0(pathToFigures,"violinPlot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_Microglia_Soussa2017.pdf"), width = 16, height = 8)
VlnPlot(object   = integratedData_choosenDim,
        features = symbols_conversion[symbols_conversion$MGI.symbol == cellType_markers_Microglia_Soussa2017$Gene,"HGNC.symbol"]
        ) +
    #ggtitle(label = "SMARCB1") +
    xlab("\nCluster") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 20),
          plot.title = element_text(hjust = 0.5, face = "bold")
          )
#dev.off()

```


#### Microglia gene set from Zhang2014

```{r immuneCells}

head(cellType_markers)

# Select immune cells canonical markers
cellType_markers_Microglia_Zhang2014 <- filter(.data = cellType_markers,
                                               Source == "Zhang et al., 2014",
                                               Cell.type == "Microglia",
                                               Organism == "murine"
                                               )


length(cellType_markers_Microglia_Zhang2014$Gene)

cluster <- vector()
k <- vector()
K <- vector()
pVal <- vector()
for (clust in levels(integratedData_choosenDim.markers$cluster))
{
    g <- filter(.data    = integratedData_choosenDim.markers_plus,
                cluster  == clust
                )$MGI.symbol
    K <- c(K,length(g))
    G <- as.vector(cellType_markers_Microglia_Zhang2014$Gene)
    # Intersect cluster gene markers and cell type markers
    gG <- intersect(g,G)
    k <- c(k,length(gG))
    # Hypergeometric test
    p <- computeHypergeometrix_pval(x = g,
                                    y = G,
                                    N = nrow(integratedData_choosenDim)
                                    )
    pVal <- c(pVal,p)
    cluster <- c(cluster,clust)
}
overlap_integratedData_markers_Microglia_Zhang2014 <- data.frame(cluster,k,K,pVal)

# Compute q values
overlap_integratedData_markers_Microglia_Zhang2014$qVal <- qvalue(p = overlap_integratedData_markers_Microglia_Zhang2014$pVal)$qvalues

overlap_integratedData_markers_Microglia_Zhang2014

# Keep cluster order for ggplot
overlap_integratedData_markers_Microglia_Zhang2014$cluster <- factor(overlap_integratedData_markers_Microglia_Zhang2014$cluster,
                                                                     levels = overlap_integratedData_markers_Microglia_Zhang2014$cluster
                                                                     )

# Plot barplot
#pdf(file = paste0(pathToFigures,"barplot_scRNAseq_integratedData_ATRT_SHH_VermisAnterior_markers_Microglia_Zhang2014.pdf"), width = 16, height = 10)
ggplot(data    = overlap_integratedData_markers_Microglia_Zhang2014,
       mapping = aes(x = cluster,
                     y = -log10(qVal)
                     )
       ) +
    geom_bar(stat = "identity",
             fill = "brown"
             ) +
    #coord_flip() +
    xlab("\nCluster") +
    ylab("-log10(q-value)\n") +
    #ylim(0,xlim) +
    #geom_text(mapping = aes(label=paste0("(",k,"/",K,")")),
     #         size = 10,
      #        hjust = -0.1
       #       ) +
    theme_bw() +
    theme(text         = element_text(size=40, color = "black"),
          panel.border = element_blank(),
          axis.text.x  = element_text(color = "black")
          )
#dev.off()

```


# Session info

```{r sessionInfo}

# Session info
sessionInfo()

```
